---
title: "2 - Your data budget - Classwork"
subtitle: "Survival analysis with tidymodels"
editor_options: 
  chunk_output_type: console
---

We recommend restarting R between each slide deck!

## Data on shelter cats

```{r}
library(tidymodels)
library(censored)
library(modeldata)

cat_adoption

cat_adoption <- cat_adoption %>% 
  mutate(event_time = Surv(time, event), .keep = "unused", .before = everything()) 
```

## Your turn

When is a good time to split your data?

## Data splitting and spending

```{r}
set.seed(123)

cat_split <- initial_split(cat_adoption)
cat_split
```

Extract the training and testing sets

```{r}
cat_train <- training(cat_split)
cat_test <- testing(cat_split)
```

## Your turn

Split your data so 20% is held out for the test set.

Try out different values in `set.seed()` to see how the results change.

Hint: Which argument in `initial_split()` handles the proportion split into training vs testing?

```{r}
# Your code here!

```

## Your turn

Explore the `cat_train` data on your own!

- What does the Kaplan-Meier curve look like for the outcome, `event_time`?
- How does `event_time` differ across the categorical variables?
- What's the distribution of the location information, `latitude` and `longitude`?

Extension/Challenge: How would you visualize the relelationship between the outcome and longitude and latitude, respectively?

```{r}
# Your code here!

```
